---
title: "Working with dates"
author: "Erik Bulow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rccdates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

Registers managed by the Swedish Cancer centers (quality registers and the cancer register) can have date variables in different formats. This is most evident for old registers handled by Rockan. 

## The `as.Dates` function

The function `as.Dates` (plural) is intended to replace the ordinary `as.Date` (singular) function from base R. 


## Missing dates

RCC dates are usually in the form `%Y-%m-%d`, such as `r Sys.Date()`. These are recognised by ordinary R-functions such as `as.Date` if there are no missing values or if missing values are coded as `NA`. It is however common for RCC that missing dates are coded as empty strings:

```{r, error = TRUE, purl = FALSE}
suppressPackageStartupMessages(library(incadata))

d <- c("", "2016-06-17")
as.Date(d)
```

`as.Dates` on the other hand has no problem:

```{r}
as.Dates(d)
```



## Missing day and/or month

Date from old Rockan registers can also have missing days and months (coded as "00"). `as.Dates` will recognise these dates and will replace "00" by an approximate date:

```{r, error = TRUE, purl = FALSE}
as.Date("2000-01-00")  # as.Date fails!
as.Dates("2000-01-00") # Missing day
as.Dates("2000-00-00") # Missing month and day
```


## Year and week

Some old dates can have format `%Y%V` (see `?strptime`), such as "7403" for week 3 in 1974. This is tricky for four reasons:

* Exact date is unknown but has to be approximated.
* Different countries numerate weeks of year differently. Sweden use ISO-8601 ("If the week [starting on Monday] containing 1 January has four or more days in the new year, then it is considered week 1. Otherwise, it is the last week of the previous year, and the next week is week 1.") but R does not.
* Procedures for week number also differs between operating systems.

```{r, error = TRUE, purl = FALSE}
as.Date("7403")
as.Dates("7403")
```


It is also possible to have a mixture of different dates within the same vector:
```{r}
as.Dates(c("", NA, "2000-01-01", "20000101", "20000000", "7403"))
```

