---
title: "rccdates"
author: "Erik Bulow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rccdates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Intro

Registers managed by the Swedish Cancer centers (quality registers and the cancer register) have date variables in different formats. This package can recognise and handle such dates.

```{r}
library(incadata)
```


# Ordinary dates


RCC dates are usually in the form `%Y-%m-%d`, such as "2016-06-17". These are recognised by ordinary R-functions such as `as.Date` if there are no missing values or if missing values are coded as NA. It is however common for RCC that missing dates are coded as empty strings. Then:

```{r, error = TRUE, purl = FALSE}
d <- c("", "2016-06-17")
as.Date(d)
```

The `as.Dates` function (note the plural) might then be easier to use.

```{r}
as.Dates(d)
```



# Non standard dates

The oringinal motivation for the package was to handle old date variables from the cancer register. Days and even months are sometimes coded as "00" (unknown). If so happens, `as.Dates` (note the plural) might still recognise the date and will replace "00" by an approximate date:

```{r, error = TRUE, purl = FALSE}
as.Date("2000-01-00") # as.Date fails!
as.Dates("2000-01-00") # Missing day
as.Dates("2000-00-00") # Missing month and day
```


Some old dates might also be in the format `%Y%V` (see `?strptime`), such as "7403" for week 3 in 1974. This is tricky for four reasons:

* Exact date is unknown but has to be approximated
* Different countries numerate weeks of year differently. Sweden use ISO-8601 ("If the week [starting on Monday] containing 1 January has four or more days in the new year, then it is considered week 1. Otherwise, it is the last week of the previous year, and the next week is week 1.") but R does not.
* Procedures for week number differ also between different operating systems

```{r, error = TRUE, purl = FALSE}
as.Date("7403")
as.Dates("7403")
```


It is also possible to have a mixture of different dates within the same vector:
```{r}
as.Dates(c("", NA, "2000-01-01", "20000101", "20000000", "7403"))
```

